---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo = TRUE}
setwd("~/Documents/Project/RepData_PeerAssessment1")
library(ggplot2)
unzip("./activity.zip")

act <- read.csv("./activity.csv")
act$date <- as.POSIXct(act$date, "%Y-%m-%d")
weekday <- weekdays(act$date)
act <- cbind(act,weekday)

summary(act)

## What is mean total number of steps taken per day?
```{r, echo = TRUE}
act_total_steps <- with(act, aggregate(steps, by = list(date), FUN = sum, 
na.rm = TRUE))
names(act_total_steps) <- c("date", "steps")
hist(act_total_steps$steps, main = "Total number of steps taken per day", 
xlab = "Total steps taken per day", col = "green")

mean(act_total_steps$steps)
median(act_total_steps$steps)

## What is the average daily activity pattern?
```{r, echo = TRUE}
average_daily_act <- aggregate(act$steps, by=list(act$interval), FUN=mean, 
na.rm=TRUE)
names(average_daily_act) <- c("interval", "mean")
plot(average_daily_act$interval, average_daily_act$mean, type = "l", 
col="green", xlab="Interval", ylab="Average number of steps", 
main="Average number of steps per intervals")

average_daily_act[which.max(average_daily_act$mean), ]$interval

## Imputing missing values
```{r, echo = TRUE}
sum(is.na(act$steps))
imputed_steps <- average_daily_act$mean[match(act$interval, 
average_daily_act$interval)]
act_imputed <- transform(act, steps = ifelse(is.na(act$steps), 
yes = imputed_steps, no = act$steps))
total_steps_imputed <- aggregate(steps ~ date, act_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per 
day", main = "Total number of steps taken each day")

mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}
act$date <- as.Date(strptime(act$date, format="%Y-%m-%d"))
act$datetype <- sapply(act$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
act_by_date <- aggregate(steps~interval + datetype, act, mean, na.rm = TRUE)
plot<- ggplot(act_by_date, aes(x = interval , y = steps, color = datetype)) +
      geom_line() +
      labs(title = "Average daily steps by type of date", x = "Interval", 
      y = "Average number of steps") +
      facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)


